---
title: "Basic SIR Modeling"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

## Modelling an infected cohort
Loading the packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(deSolve) 
library(reshape2)
library(ggplot2)  
```

Filling in the input and specifying timesteps:

```{r}
initial_number_infected <- 1000000  

initial_number_recovered <- 0      
recovery_rate <- 10^-1               
follow_up_duration <- 4*7          

initial_state_values <- c(I = initial_number_infected, 
                          R = initial_number_recovered)  

parameters <- c(gamma = recovery_rate)  

times <- seq(from = 0, to = follow_up_duration, by = 1) 
```

The model function

```{r}
cohort_model <- function(time, state, parameters) {    
    
    with(as.list(c(state, parameters)), {  
      dI <- -gamma * I
      dR <- gamma * I
        
    return(list(c(dI, dR)))                             
    })
}
```

Solving the model equations using the ode() function in the deSolve package and saving them as a dataframe object:

```{r}
output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = cohort_model,
                            parms = parameters))
```

Based on the output, how many people have recovered after 4 weeks
```{r}
output[output$time == 28, c("time","R")] 
```

Ploting the Output

```{r}
output_long <- melt(as.data.frame(output), id = "time")                  


ggplot(data = output_long,         
       aes(x = time, y = value, colour = variable, group = variable)) + 
  geom_line() +                                                          
  xlab("Time (days)")+                          
  ylab("Number of people") +                  
  labs(title = paste("Number infected and recovered over time when gamma =",
               parameters["gamma"],"days^-1")) 
```

Based on the plot, at what timepoint were infected and recovered individuals equal in number
```{r}
output[output$time == 7,]
```

Varying  ð›¾ 
Average duration of infection = 2 days so the recovery rate = 1/2 = 0.5 days âˆ’1
```{r}
parameters <- c(gamma = 0.5)

# Solving the model
output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = cohort_model,
                            parms = parameters))

# Plotting the output
output_long <- melt(as.data.frame(output), id = "time")   
# turn output dataset into long format

ggplot(data = output_long,       
       aes(x = time, 
           y = value, 
           colour = variable, 
           group = variable)) +  
  geom_line() +                                
  xlab("Time (days)")+                          
  ylab("Number of people") +                    
  labs(title = paste("Number infected and recovered over time when gamma =",
                      parameters["gamma"],"days^-1")) + # add title
  scale_color_brewer(palette = "Set1")
```

Average duration of infection = 20 days so the recovery rate = 1/20 = 0.05 days âˆ’1
```{r}
parameters <- c(gamma = 0.05)

output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = cohort_model,
                            parms = parameters))

output_long <- melt(as.data.frame(output), id = "time") 

ggplot(data = output_long,                                 
       aes(x = time, 
           y = value, 
           colour = variable, 
           group = variable)) +  
  geom_line() +                  
  xlab("Time (days)")+          
  ylab("Number of people") +   
  labs(title = paste("Number infected and recovered over time when gamma =",
                    parameters["gamma"],"days^-1"))
```

##Simulating competing hazards

Question: what do  ð›¾  (gamma) and  ðœ‡  (mu) represent? Draw the model diagram corresponding to these equations.
Now, incorporate the new compartment and transition into your code. Since we already have a working scaffold for a model from the first coding etivity, it is good practice to copy and adapt it. In the cell below, copy-paste the model function from the previous etivity and modify it according to the new differential equations above. Make sure you understand what every line of the code means, and run it once you're ready
```{r}
library(deSolve)
library(reshape2)
library(ggplot2)

cohort_model <- function(time, state, parameters) {  

    with(as.list(c(state, parameters)), {  
                                           
        
    # The differential equations
      dI <- -gamma * I - mu * I  
      dR <- gamma * I          
      dM <- mu * I              
                                 
   
    return(list(c(dI, dR, dM))) 
    })}
```

Question: after 4 weeks, do you expect more people to have recovered or more people to have died, and why? Proceed with the next steps to check if you are right.

```{r}
initial_state_values <- c(I = 1000000, R =0, M=0 )
parameters <- c(gamma = 0.1, mu =0.2)
times <- seq(from= 0, to= 4*7, by =1)
output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = cohort_model,
                            parms = parameters))
```

```{r}
output_long <- melt(as.data.frame(output), id = "time")
ggplot(output_long, aes(time, value, color = variable, group = variable))+
geom_line()+
xlab("Time (days)")+
ylab("Number of people")+
labs(color = "Compartment")
```

Question: based on the model output, what proportion of the initially infected cohort died before recovering over the 4 week period?

```{r}
output
output[output$time == 28, ]
output[29, "M"]/1000000
```
Question: now use the competing hazards formula given in the video lecture to calculate the case fatality rate. Does this agree with your answer to the previous question?
```{r}
parameters["mu"]/(parameters["mu"]+parameters["gamma"])
```
Question: Which value of  ðœ‡  do you need to get a case fatality rate of 50% assuming  ð›¾  stays fixed?
```{r}
parameters <- c(gamma = 0.1, mu = 0.1)

output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = cohort_model,
                            parms = parameters))

output[29,"M"]/1000000

```

## MODEL INPUTS:
```{r}
initial_state_values <- c(S = 999999, I = 1, R = 0)

parameters <- c(lambda = 0.2, gamma = 0.1)
```

## TIMESTEPS:
```{r}
times <- seq(from = 0, to = 60, by = 1)
```

## SIR MODEL FUNCTION 
```{r}
sir_model <- function(time, state, parameters){
                        with(as.list(c(state, parameters)),{
                        
                        dS <- -lambda*S
                        dI <-  lambda*S-gamma*I
                        dR <- gamma*I
                        
                        return(list(c(dS, dI, dR)))
                        })
}

```

# MODEL OUTPUT (solving the differential equations):
```{r}
output <- as.data.frame(ode(y= initial_state_values, times= times, func=sir_model, parms = parameters))

output
```

## Plot:
```{r}
output_long <- melt(as.data.frame(output), id = "time")                  

ggplot(data = output_long,                                               
       aes(x = time, y = value, color = variable, group = variable)) +  
  geom_line() +                                                          
  xlab("Time (days)")+                                                   
  ylab("Number of people") +                                          
  labs(color = "Compartment") 
```
