---
title: "SIR Model with a Dynamic Force of Infection"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Load Packages

```{r}
library(deSolve)
library(reshape2)
library(ggplot2)
```

## Model Inputs 

```{r}
initial_state_values <- c(S = 999999, I = 1, R= 0)
parameters <- c(beta= 1, gamma = 0.1)
```

## Time Steps

```{r}
times <- seq(from = 0, to = 60, by = 1)
```

## SIR Model Function

```{r}
sir_model <- function(time, state, parameters){
      with(as.list(c(state, parameters)), {
        N<- S+I+R
        lambda <- beta*I/N
        dS <- -lambda * S
        dI <- lambda * S -gamma * I
        dR <- gamma * I
      return(list(c(dS, dI, dR)))   
      })
}
```

## Output

```{r}
output <- as.data.frame(ode( y= initial_state_values, times = times, func = sir_model,
                             parms = parameters))
```

## Plot

```{r}
output_long <- melt(as.data.frame(output), id = "time")
ggplot(output_long, aes(x= time, y= value, color = variable, group = variable))+
  geom_line()+
  xlab("Time (days)")+
  ylab("Number of people")+
  labs(color = "Compartment")
```

What is the peak prevalence?

```{r}
output[output$time == 19,]
```

## SIR Dynamics with Varying Parameters

## Model Inputs
```{r}
initial_state_values <- c(S = 1000000-1,  
                          I = 1,          
                          R = 0)         


parameters <- c(beta = 0.5,     
                gamma = 0.25) 
```

## TIMESTEPS:
```{r}
times <- seq(from = 0, to = 100, by = 1)  
```

## SIR MODEL FUNCTION:

```{r}
sir_model <- function(time, state, parameters) {  

    with(as.list(c(state, parameters)), {   # tell R to unpack variable names from the state and parameters inputs    
        
    # Total population size N 
      N <- S+I+R
      
    
      lambda <- beta * I/N
        
    ## Differential equations
      dS <- -lambda * S              
      dI <- lambda * S - gamma * I    
                                      
      dR <- gamma * I                 
    return(list(c(dS, dI, dR))) 
    })
  
}

```

## MODEL OUTPUT:
```{r}
output <- as.data.frame(ode(y = initial_state_values, 
                            times = times, 
                            func = sir_model,
                            parms = parameters))
```

## PLOT THE OUTPUT
```{r}
output_long <- melt(as.data.frame(output), id = "time")                  

output_long$proportion <- output_long$value/sum(initial_state_values)

```

## Plot the Final Output
```{r}
ggplot(data = output_long,                                               
       aes(x = time, y = proportion, color = variable, group = variable)) +  
  geom_line() +                                                          
  xlab("Time (days)")+                                                  
  ylab("Proportion of the population") +                                
  labs(color = "Compartment") 
```
## The Parameters
```{r}
parameters <- c(beta = 0.1,     
                gamma = 0.25) 
```

## The Differential Equations

```{r}
output <- as.data.frame(ode(y = initial_state_values, times = times, 
                            func = sir_model, parms = parameters))
```

## Ploting the Output
```{r}
output_long <- melt(as.data.frame(output), id = "time")                  


output_long$proportion <- output_long$value/sum(initial_state_values)

ggplot(data = output_long,                                               
       aes(x = time, y = proportion, colour = variable, group = variable)) +  
  geom_line() +                                                          
  xlab("Time (days)")+                                                  
  ylab("Proportion of the population") +                                
  labs(colour = "Compartment")  
```
## Simulating the Effective Reproduction Number Reff

```{r}
# LOAD THE PACKAGES:
library(deSolve)
library(reshape2)
library(ggplot2)

# MODEL INPUTS:

initial_state_values <- c(S = 1000000-1,   
                          I = 1,          
                          R = 0)           


parameters <- c(beta = 0.4,      
                gamma = 0.1)     

# TIMESTEPS:

times <- seq(from = 0, to = 100, by = 1)   

# SIR MODEL FUNCTION: 

# The model function 
sir_model <- function(time, state, parameters) {  

    with(as.list(c(state, parameters)), {  
        
    # Total population size N 
      N <- S+I+R
      
    # Defining lambda as a function of beta and I:
      lambda <- beta * I/N
        
    # The differential equations
      dS <- -lambda * S               
      dI <- lambda * S - gamma * I     
                                      
      dR <- gamma * I                 
      
    
    return(list(c(dS, dI, dR))) 
    })
  
}

# Model Output:

output <- as.data.frame(ode(y = initial_state_values, times = times, func = sir_model, parms = parameters))
```

```{r}
# LOAD THE PACKAGES:
library(deSolve)
library(reshape2)
library(ggplot2)

```

# MODEL INPUTS:
```{r}
initial_state_values <- c(S = 1000000-1,   
                          I = 1,          
                          R = 0)           


parameters <- c(beta = 0.4,      
                gamma = 0.1)     
```
# TIMESTEPS:
```{r}
times <- seq(from = 0, to = 100, by = 1)   
```
# SIR MODEL FUNCTION: 
```{r}
# The model function 
sir_model <- function(time, state, parameters) {  

    with(as.list(c(state, parameters)), {  
        
    # Total population size N 
      N <- S+I+R
      
    # Defining lambda as a function of beta and I:
      lambda <- beta * I/N
        
    # The differential equations
      dS <- -lambda * S               
      dI <- lambda * S - gamma * I     
                                      
      dR <- gamma * I                 
      
    
    return(list(c(dS, dI, dR))) 
    })
  
}
```
# Model Output:
```{r}
output <- as.data.frame(ode(y = initial_state_values, times = times, func = sir_model, parms = parameters))
```

```{r}
output_long <- melt(as.data.frame(output), id = "time")                  

# Proportion in each compartment as a column in the long-format output
output_long$proportion <- output_long$value/sum(initial_state_values)

```

# Plot 
```{r}
ggplot(data = output_long,                                               
       aes(x = time, y = proportion, colour = variable, group = variable)) +  
  geom_line() +                                                          
  xlab("Time (days)")+                                                   
  ylab("Proportion of the population") +                                 
  labs(colour = "Compartment",                                            
       title = "Proportion susceptible, infected and recovered over time") +                                                               
  theme(legend.position = "bottom")                                      
```
# Calculating the effective reproduction number in a new column
```{r}
output$reff <- parameters["beta"]/parameters["gamma"] *                 
                output$S/(output$S+output$I+output$R)                    
```
# Plot Reff
```{r}
ggplot(data = output,                                                    
       aes(x = time, y = reff)) +                                        
  geom_line() +                                                          
  xlab("Time (days)")+                                                   
  ylab("Reff") +                                                         
  labs(title = "Effective reproduction number over time")                
```

